---
title: "Data Manipulation in R with dplyr"
author: "Alan"
date: "25 de agosto de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Course Description

_In this interactive tutorial, you will learn how to perform sophisticated dplyr techniques to carry out your data manipulation with R. First you will master the five verbs of R data manipulation with dplyr: select, mutate, filter, arrange and summarise. Next, you will learn how you can chain your dplyr operations using the pipe operator of the magrittr package. In the final section, the focus is on practicing how to subset your data using the group_by function, and how you can access data stored outside of R in a database. All said and done, you will be familiar with data manipulation tools and techniques that will allow you to efficiently manipulate data._

### Introduction to dplyr and tbls

_Introduction to the dplyr package and the tbl class. Learn the philosophy that guides dplyr, discover some useful applications of the dplyr package, and meet the data structures that dplyr uses behind the scenes._

### Load the dplyr and hflights package

Welcome to the interactive exercises part of your dplyr course. Here you will learn the ins and outs of working with dplyr. dplyr is an R package, a collection of functions and data sets that enhance the R language.

Throughout this course you will use dplyr to analyze a data set of airline flight data containing flights that departed from Houston. This data is stored in a package called hflights.

Both dplyr and hflights are already installed on DataCamp's servers, so loading them with library() will get you up and running.

### Instructions
Load the dplyr package.
Load the hflights package. A variable called hflights will become available, a data.frame representing the data set.
Use both head() and summary() on the hflights data frame to get to know the data. Can you guess the meaning of all variables?


```{r}
#Install package

#install.packages("dplyr")
#install.packages("hflights")

# Load the dplyr package
library("dplyr")

# Load the hflights package
library("hflights")

# Call both head() and summary() on hflights
head(hflights)
summary(hflights)

```

###Explore the data set

A data scientist must be familiar with his or her data. Experiment with the data set in the console and maybe try to generate some insightful plots. For your convenience, hflights is already loaded into the workspace.

How many observations and how many variables are contained in the hflights data set?

```{r}

dim(hflights)
glimpse(hflights)

```

### Convert data.frame to tibble

As Garrett explained, a tbl (pronounced tibble) is just a special kind of data.frame. They make your data easier to look at, but also easier to work with. On top of this, it is straightforward to derive a tbl from a data.frame structure using tbl_df().

The tbl format changes how R displays your data, but it does not change the data's underlying data structure. A tbl inherits the original class of its input, in this case, a data.frame. This means that you can still manipulate the tbl as if it were a data.frame. In other words, you can do anything with the hflights tbl that you could do with the hflights data.frame.

Instructions
Convert hflights (which is a data.frame) into a tbl, also named hflights.
Display the new hflights in your console window. Notice the easy-to-read layout.
To see how tbls behave like data.frames, save the UniqueCarrier column of hflights as an object named carriers, using base R syntax only.

```{r}
# Both the dplyr and hflights packages are loaded

# Convert the hflights data.frame into a hflights tbl
hflights <- tbl_df(hflights)

# Display the hflights tbl
hflights

# Create the object carriers
carriers <- hflights$UniqueCarrier

```

###Changing labels of hflights, part 1 of 2

A bit of cleaning would be a good idea since the UniqueCarrier variable of hflights uses a confusing code system.

To do this, let's work with a lookup table, that comes in the form of a named vector. When you subset the lookup table with a character string (like the character strings in UniqueCarrier), R will return the values of the lookup table that correspond to the names in the character string. To see how this works, run following code in the console:

two <- c("AA", "AS")
lut <- c("AA" = "American", 
         "AS" = "Alaska", 
         "B6" = "JetBlue")
two <- lut[two]
two

Instructions
Add a new Carrier column to hflights by combining lut with the UniqueCarrier column of hflights.
It's rather hard to see if you did things right, since the Carrier variable does not appear when you print hflights. Use the glimpse() function on hflights instead.

```{r}

# Both the dplyr and hflights packages are loaded into workspace
lut <- c("AA" = "American", "AS" = "Alaska", "B6" = "JetBlue", "CO" = "Continental", 
         "DL" = "Delta", "OO" = "SkyWest", "UA" = "United", "US" = "US_Airways", 
         "WN" = "Southwest", "EV" = "Atlantic_Southeast", "F9" = "Frontier", 
         "FL" = "AirTran", "MQ" = "American_Eagle", "XE" = "ExpressJet", "YV" = "Mesa")

# Add the Carrier column to hflights
hflights$Carrier <- lut[hflights$UniqueCarrier]

# Glimpse at hflights
glimpse(hflights)


```

###Changing labels of hflights, part 1 of 2

A bit of cleaning would be a good idea since the UniqueCarrier variable of hflights uses a confusing code system.

To do this, let's work with a lookup table, that comes in the form of a named vector. When you subset the lookup table with a character string (like the character strings in UniqueCarrier), R will return the values of the lookup table that correspond to the names in the character string. To see how this works, run following code in the console:

two <- c("AA", "AS")
lut <- c("AA" = "American", 
         "AS" = "Alaska", 
         "B6" = "JetBlue")
two <- lut[two]
two
Instructions
Add a new Carrier column to hflights by combining lut with the UniqueCarrier column of hflights.
It's rather hard to see if you did things right, since the Carrier variable does not appear when you print hflights. Use the glimpse() function on hflights instead.


```{r}
# Both the dplyr and hflights packages are loaded into workspace
lut <- c("AA" = "American", "AS" = "Alaska", "B6" = "JetBlue", "CO" = "Continental", 
         "DL" = "Delta", "OO" = "SkyWest", "UA" = "United", "US" = "US_Airways", 
         "WN" = "Southwest", "EV" = "Atlantic_Southeast", "F9" = "Frontier", 
         "FL" = "AirTran", "MQ" = "American_Eagle", "XE" = "ExpressJet", "YV" = "Mesa")

# Add the Carrier column to hflights
hflights$Carrier <- lut[hflights$UniqueCarrier]

# Glimpse at hflights
glimpse(hflights)


```


###Changing labels of hflights, part 2 of 2

Let's try a similar thing, but this time to change the labels in the CancellationCode column. This column lists reasons why a flight was cancelled using a non-informative alphabetical code. Execute

unique(hflights$CancellationCode)
A lookup table lut has already been created for you, that converts the alphabetical codes into more meaningful strings.

Instructions
Use lut to change the labels of the CancellationCode column of hflights. Store the recoded vector in a new column Code.
As before, check your results by glimpsing at them.


```{r}

# The hflights tbl you built in the previous exercise is available in the workspace.

# The lookup table
lut <- c("A" = "carrier", "B" = "weather", "C" = "FFA", "D" = "security", "E" = "not cancelled")

# Add the Code column
hflights$Code <- lut[hflights$CancellationCode]

# Glimpse at hflights
glimpse(hflights)


```

# Select and mutate

_Get familiar with dplyr's manipulation verbs. Meet the five verbs and then practice using the mutate and select verbs._



Comparison to base R

To see the added value of the dplyr package, it is useful to compare its syntax with base R. Up to now, you have only considered functionality that is also available without the use of dplyr. The elegance and ease-of-use of dplyr is a great plus though.

Instructions
Finish the select() calls to match the results of the base R commands. Try to make your calls as concise as possible.

```{r}
# both hflights and dplyr are available

# Finish select call so that ex1d matches ex1r
ex1r <- hflights[c("TaxiIn", "TaxiOut", "Distance")]
ex1d <- select(hflights, starts_with("TaxI"), Distance)
ex1d
# Finish select call so that ex2d matches ex2r
ex2r <- hflights[c("Year", "Month", "DayOfWeek", "DepTime", "ArrTime")]
ex2d <- select(hflights, Year:DayOfWeek, -3, DepTime, ArrTime)
ex2d
# Finish select call so that ex3d matches ex3r
ex3r <- hflights[c("TailNum", "TaxiIn", "TaxiOut")]
ex3d <- select(hflights, starts_with("Ta"))
ex3d
```



### Choosing is not losing! The select verb

To answer the simple question whether flight delays tend to shrink or grow during a flight, we can safely discard a lot of the variables of each flight. To select only the ones that matter, we can use select().
As an example, take the following call, that selects the variables var1 and var2 from the data frame df.
select(df, var1, var2)
You can also use : to select a range of variables and - to exclude some variables, similar to indexing a data.frame with square brackets. You can use both variable's names as well as integer indexes. This call selects the four first variables except for the second one of a data frame df:
select(df, 1:4, -2)
select() does not change the data frame it is called on; you have to explicitly assign the result of select() to a variable to store the result.

Instructions

	* Use select() to print out a tbl that contains only the columns ActualElapsedTime, AirTime, ArrDelay and DepDelay of hflights.
	* Print out a tbl with the columns Origin up to and including Cancelledof hflights.
	* Find the most concise way to select: columns Year up to and including DayOfWeek, columns ArrDelay up to and including Diverted. You can examine the order of the variables in hflights with names(hflights) in the console.

```{r}
# hflights is pre-loaded as a tbl, together with the necessary libraries.

# Print out a tbl with the four columns of hflights related to delay
select(hflights, ActualElapsedTime, AirTime, ArrDelay, DepDelay)

# Print out the columns Origin up to Cancelled of hflights
select(hflights, Origin:Cancelled)

# Answer to last question: be concise!
select(hflights, Year:DayOfWeek, ArrDelay:Diverted)



```


### Helper functions for variable selection

dplyr comes with a set of helper functions that can help you select groups of variables inside a select() call:

starts_with("X"): every name that starts with "X",
ends_with("X"): every name that ends with "X",
contains("X"): every name that contains "X",
matches("X"): every name that matches "X", where "X" can be a regular expression,
num_range("x", 1:5): the variables named x01, x02, x03, x04 and x05,
one_of(x): every name that appears in x, which should be a character vector.
Pay attention here: When you refer to columns directly inside select(), you don't use quotes. If you use the helper functions, you do use quotes.

Instructions
Use select() and a helper function to print out a tbl that contains just ArrDelay and DepDelay of hflights.
Use a combination of helper functions and variable names to print out only the UniqueCarrier, FlightNum, TailNum, Cancelled, and CancellationCode columns of hflights.
Find the most concise way to return the following columns with select and its helper functions: DepTime, ArrTime, ActualElapsedTime, AirTime, ArrDelay, DepDelay. Use only helper functions!

```{r}
# As usual, hflights is pre-loaded as a tbl, together with the necessary libraries.

# Print out a tbl containing just ArrDelay and DepDelay
select(hflights, ends_with("Delay"))

# Print out a tbl as described in the second instruction, using both helper functions and variable names
select(hflights, UniqueCarrier, contains("Cancell"), ends_with("Num"))

# Print out a tbl as described in the third instruction, using only helper functions.
select(hflights, ends_with("Delay"), ends_with("Time")) 


```


# hflights and dplyr are loaded and ready to serve you.

# Add the new variable ActualGroundTime to a copy of hflights and save the result as g1.
g1 <- mutate(hflights, ActualGroundTime = ActualElapsedTime - AirTime)

# Add the new variable GroundTime to g1. Save the result as g2.
g2 <- mutate(g1, GroundTime = TaxiIn + TaxiOut)

# Add the new variable AverageSpeed to g2. Save the result as g3.
g3 <- mutate(g2, AverageSpeed = Distance / AirTime * 60)

# Print out g3



```{r}
# hflights and dplyr are loaded and ready to serve you.

# Add the new variable ActualGroundTime to a copy of hflights and save the result as g1.
g1 <- mutate(hflights, ActualGroundTime = ActualElapsedTime - AirTime)

# Add the new variable GroundTime to g1. Save the result as g2.
g2 <- mutate(g1, GroundTime = TaxiIn + TaxiOut)

# Add the new variable AverageSpeed to g2. Save the result as g3.
g3 <- mutate(g2, AverageSpeed = Distance / AirTime * 60)

# Print out g3
g3
```



#Add multiple variables using mutate

So far you've added variables to hflights one at a time, but you can also use mutate() to add multiple variables at once. To create more than one variable, place a comma between each variable that you define inside mutate().

mutate() even allows you to use a new variable while creating a next variable in the same call. In this example, the new variable x is directly reused to create the new variable y:

mutate(my_df, x = a + b, y = x + c)
Instructions
Adapt the code that builds m1: add a variable loss_ratio, which is the ratio of loss to DepDelay.
Create a tbl m2 from hflights by adding three variables:
TotalTaxi, which is the sum of TaxiIn and TaxiOut;
ActualGroundTime, which is the difference of ActualElapsedTime and AirTime;
Diff, the difference between the two newly created variables. This column should be zero for all observations!

```{r}
# hflights and dplyr are ready, are you?

# Add a second variable loss_ratio to the dataset: m1
m1 <- mutate(hflights, loss = ArrDelay - DepDelay, loss_ratio = loss / DepDelay)
m1
# Add the three variables as described in the third instruction: m2
m2 <- mutate(hflights, TotalTaxi = TaxiIn + TaxiOut, ActualGroundTime = ActualElapsedTime - AirTime, Diff = TotalTaxi - ActualGroundTime)
m2
```


















# Filter and arrange

_Learn how to search through the observations in your data set (and extract useful observations) with the filter function. Rearrange the observations in your data set with the arrange verb._

# Summarise and the pipe operator

_Master the data manipulation verb summarise, and practice combining the five verbs to solve advanced data manipulation tasks. Learn to chain the operators together with the piping operator._


# Group_by and working with databases

_Complete your mastery of data manipulation with group-wise operations and databases. Learn to use group_by to group your data into subsets of observations, and use dplyr to access data stored outside of R in a database._



```{r pressure, echo=FALSE}
plot(pressure)
```




Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
